Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-02-27 index.qmd
added aian

- Homophily not the result of voluntary action; homophily programmed into algorithms that create connections
  - Match vertically ranked consumers to vertically ranked producers
- @RN4982: like colleges compete for like students (SES)
<p class="quote" style="font-size: 0.85em;">Recommend colleges target Geomarkets, schools targeted by peer colleges</p>
<p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged. Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good. The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice [@RN4982, p. 44].</p>

@@

>>>>

- Homophily not the result of voluntary action; homophily programmed into algorithms that create connections
  - Match vertically ranked consumers to vertically ranked producers
- @RN4982: like colleges compete for like students (SES)
<p class="quote" style="font-size: 0.85em;">Recommend colleges target Geomarkets, schools targeted by peer colleges</p>
<p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main [Geo]market areas remain forgotten and hence, unchallenged. Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good. The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice [@RN4982, p. 44].</p>

@@

############################################################################
# 5) AI/AN results loop
############################################################################

writeLines("## AI/AN students")

# The five metros (with 'nyny' last) and the single Native American order
native_metros <- c("chicago", "dallas", "houston", "los_angeles", "nyny")
native_order  <- "487927"

# We want exactly 3 slides per metro, in this order:
var_rc_pairs <- list(
  c("firstgen", "row"),
  c("firstgen", "col"),
  c("race",     "row")
)

for (m in native_metros) {

  # Determine the heading text for each metro:
  # If m == 'nyny', use "New York City".
  # Otherwise, convert underscores -> spaces, then "Title Case + area".
  heading_text <- if (m == "nyny") {
    "New York City"
  } else {
    # Normal logic
    metro_no_underscores <- gsub("_", " ", m, fixed = TRUE)
    title_case           <- tools::toTitleCase(metro_no_underscores)
    paste0(title_case, " area")
  }

  # For each metro, we produce 3 slides
  for (i in seq_along(var_rc_pairs)) {

    # If this is the 1st slide => show in TOC (no .hide).
    # Otherwise => hide from TOC with {.hide}.
    if (i == 1) {
      writeLines(str_c("### ", heading_text))
    } else {
      writeLines(str_c("### ", heading_text, " {.hide}"))
    }

    # Extract the variable (firstgen/race) and approach (row/col)
    g  <- var_rc_pairs[[i]][1]
    rc <- var_rc_pairs[[i]][2]

    # Build the figure name: e.g. "race_row_plot_487927"
    graph_type_native <- str_c(g, "_", rc, "_plot_", native_order)

    # Call your helper function; it prints "#### <title from _title.txt>"
    # and displays the figure + footnotes.
    insert_figure(
      rq         = "rq2",
      metro      = m,
      graph_type = graph_type_native
    )
  }
}

### 2025-02-27 text/front_end_v2.qmd
added aian

>>>>

**ABSTRACT**

*The Structure of College Choice* [@RN4982] created "Geomarkets" and the "Market Segment Model." Geomarkets carve states and metropolitan areas into smaller geographic units, meant to define local recruiting markets. The Market Segment Model predicts how student demand for a particular college varies by Geomarket, based on the socioeconomic characteristics of households. Geomarkets became an input for two College Board products that help colleges recruit students. First, the Enrollment Planning Service (EPS) software recommends specific Geomarkets and high schools from which colleges should recruit. Second, the Student Search Service sells the contact information of prospective students -- referred to as "student lists" -- and colleges can filter by Geomarket to determine which prospect profiles they purchase. We draw from scholarship on quantification, particularly the discussions of correlation and homophily by @RN4975, to conceptualize how recruiting products incorporate Geomarkets. We address two research questions: What is the socioeconomic and racial variation between Geomarkets and how does this variation change over time? How does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets? We answer RQ1 by analyzing Census data from 1980, 2000, and 2020. We answer RQ2 using data on student lists purchased by public universities, which we collected by issuing public records requests. We utilize a quantitative case study design. Metropolitan areas are cases. Analyses consist of descriptive statistics and interactive maps.
